---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidymedia

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of **tidymedia** is to provide tools for easily working with media (e.g., image, audio, and video) files within R and the tidyverse.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jmgirard/tidymedia")
```

### Dependencies

#### MediaInfo

**tidymedia** uses [MediaInfo](https://mediaarea.net/en/MediaInfo) to query information about media files. If you would like to use these functions, you will need to install the command line interface (CLI) version of this program. Links and instructions for doing so are available [here](https://mediaarea.net/en/MediaInfo/Download). Below are instructions for several popular platforms.

**Debian/Ubuntu**

1. Enter this code into your terminal:<br />`sudo apt-get install mediainfo`

**Windows**

1. Download the appropriate CLI .zip file from:<br /><https://mediaarea.net/en/MediaInfo/Download/Windows>
2. Extract (or copy) the contents of this .zip file to a folder on your computer such as:<br />`C:/Program Files/MediaInfo`
3. Run the following code in R (changing the path to match Step 2):<br />
`tidymedia::set_mediainfo("C:/Program Files/MediaInfo/mediainfo.exe")`

**Mac**

1. Download the appropriate CLI .dmg file from:<br /><https://mediaarea.net/en/MediaInfo/Download/Mac_OS>
2. Open the .dmg file and drag the program icon to the Applications folder

#### FFmpeg
**tidymedia** uses [FFmpeg](https://ffmpeg.org/) to encode media files. If you would like to use these functions, you will need to install the command line interface (CLI) version of this program. Links and instructions for doing so are available [here](https://ffmpeg.org/download.html). Below are instructions for several popular platforms.

**Debian/Ubuntu**

1. Enter this code into your terminal:<br />`sudo apt-get install ffmpeg`

**Windows**

1. Download the latest git version (full or essentials build) from:<br /> <https://www.gyan.dev/ffmpeg/builds/>
2. Extract the contents of the downloaded [.7z](https://www.7-zip.org/) file to a folder on your computer such as:<br />`C:/Program Files/ffmpeg`
3. Run the following code in R (changing the path to match Step 2):<br />`tidymedia::set_ffmpeg("C:/Program Files/ffmpeg/bin/ffmpeg.exe")`

**macOS Homebrew Install**

1. Open the macOS Terminal or Linux shell prompt.
2. Install [Homebrew](https://brew.sh/) by entering this code into your terminal:<br />`/bin/bash -c "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/master/install.sh)"`
3. Enter this code into your terminal:<br />`brew install ffmpeg`

**macOS Manual Install**

1. Download the latest snapshot version from:<br /><https://evermeet.cx/ffmpeg/get>
2. Extract the contents of the downloaded [.7z](https://www.7-zip.org/) file
3. Drag the extracted contents to the Applications folder

## Examples

```{r, echo=FALSE}
library(tidymedia)
```

### MediaInfo

```{r}
# Interface with the mediainfo CLI
mediainfo("--Version")
```

```{r, eval=FALSE}
# Create a row tibble using a built-in template
mediainfo_template(file = "D:/example.mp4", template = "brief")
```

```{r, echo=FALSE}
knitr::kable(mediainfo_template(file = "D:/example.mp4", template = "brief"))
```

```{r, eval=FALSE}
# Create a row tibble from specific parameters
mediainfo_query(
  file = "D:/example.mp4", 
  section = "Video", 
  parameters = c("Width", "Height", "DisplayAspectRatio")
)
```
```{r, echo=FALSE}
knitr::kable(mediainfo_query(
  file = "D:/example.mp4", 
  section = "Video", 
  parameters = c("Width", "Height", "DisplayAspectRatio")
))
```

```{r}
# Lookup common parameters
get_duration(file = "D:/example.mp4", unit = "sec")
get_height(file = "D:/example.mp4")
```

### FFmpeg

```{r}
# Interface with the mediainfo CLI
ffmpeg("-version")
```
```{r}
# Create ffmpeg commands by pipeline
set_files("D:/example.mp4", "D:/example2.mp4") %>% 
  trim_duration(start_at = 1, stop_at = 5) %>% 
  set_codec("video", "libx264") %>%
  drop_streams("audio") %>% 
  apply_autocrop() %>% 
  compile_command()
```

```{r, eval=FALSE}
# Run ffmpeg commands by pipeline
set_files("D:/example.mp4", "D:/example2.mp4") %>% 
  trim_duration(start_at = 1, stop_at = 5) %>% 
  set_codec("video", "libx264") %>%
  drop_streams("audio") %>% 
  apply_autocrop() %>% 
  run_ffmpeg()
```

## Code of Conduct

Please note that the **tidymedia** project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
