---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
# Some of the chunks below run real commands, which means starting FFmpeg,
# FFprobe or MediaInfo. Each of those is evaluated only when the program it
# needs is installed, so this README knits on a machine that has none of them --
# the machine every reader of the install section above is on. The two chunks
# that only COMPILE a command carry no guard: they start nothing, and a reader
# without the binaries should still see what the builder produces.
# `tools/vignette_chunk_guards.R` is what says which chunk is which.
has_ffmpeg <- nzchar(Sys.which("ffmpeg"))
has_ffprobe <- nzchar(Sys.which("ffprobe"))
has_mediainfo <- nzchar(Sys.which("mediainfo"))
# Recorded on stderr rather than into the page: this line is what a build log
# shows for the three programs, so a build that was meant to run without them
# can be checked rather than assumed. `message()` would be captured by knitr and
# then dropped with the rest of this chunk's output.
writeLines(
  paste0("tidymedia README build, Sys.which(): ", paste0(
    c("ffmpeg", "ffprobe", "mediainfo"), "=[",
    Sys.which(c("ffmpeg", "ffprobe", "mediainfo")), "]", collapse = " ")),
  con = stderr()
)
```

# tidymedia

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/jmgirard/tidymedia/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jmgirard/tidymedia/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/jmgirard/tidymedia/graph/badge.svg)](https://app.codecov.io/gh/jmgirard/tidymedia)
<!-- badges: end -->

The goal of **tidymedia** is to provide tools for easily working with media (e.g., image, audio, and video) files within R and the tidyverse. It wraps [FFmpeg](https://ffmpeg.org/) and [MediaInfo](https://mediaarea.net/en/MediaInfo) for **reproducible media preprocessing** — batch trimming, cropping, format standardization, and metadata extraction as tibbles.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jmgirard/tidymedia")
```

### Dependencies

#### 1. MediaInfo

**tidymedia** uses [MediaInfo](https://mediaarea.net/en/MediaInfo) to query information about media files. If you would like to use these functions, you will need to install the command line interface (CLI) version of this program. Links and instructions for doing so are available [here](https://mediaarea.net/en/MediaInfo/Download). Below are instructions for several popular platforms.

**Debian/Ubuntu**

1. Enter this code into your terminal:<br />`sudo apt-get install mediainfo`
2. Check that R can find it:<br />
`tidymedia::program_status()`<br />
The `mediainfo` row should show a location and a version. If its location is
`NA`, tidymedia cannot see the program: give it the path yourself with
`tidymedia::set_mediainfo("/path/to/mediainfo")`.

**Windows**

1. Download the appropriate CLI .zip file from:<br /><https://mediaarea.net/en/MediaInfo/Download/Windows>
2. Extract (or copy) the contents of this .zip file to a folder on your computer such as:<br />`C:/Program Files/MediaInfo`
3. Run the following code in R (changing the path to match Step 2):<br />
`tidymedia::set_mediainfo("C:/Program Files/MediaInfo/mediainfo.exe")`<br />
This asks you to confirm the path before it remembers it. In a script with no
one to ask, add `confirm = FALSE`.
4. Check that R can find it:<br />
`tidymedia::program_status()`<br />
The `mediainfo` row should show a location and a version. If its location is
`NA`, the path in Step 3 is not where `mediainfo.exe` actually landed: look in
the folder from Step 2 and run `tidymedia::set_mediainfo()` again with the path
you find there.

**Mac**

1. Download the appropriate CLI .dmg file from:<br /><https://mediaarea.net/en/MediaInfo/Download/Mac_OS>
2. Open the .dmg file and drag the program icon to the Applications folder
3. The Applications folder is not on the `PATH`, so tell tidymedia where the
   program landed (changing the path to match Step 2):<br />
`tidymedia::set_mediainfo("/Applications/mediainfo")`<br />
This asks you to confirm the path before it remembers it, and it is remembered
across R sessions.
4. Check that R can find it:<br />
`tidymedia::program_status()`<br />
The `mediainfo` row should show a location and a version. If its location is
`NA`, the path in Step 3 is not where the program actually landed: find it in
the Applications folder and run `tidymedia::set_mediainfo()` again with that
path.

#### 2. FFmpeg
**tidymedia** uses [FFmpeg](https://ffmpeg.org/) to encode media files. If you would like to use these functions, you will need to install the command line interface (CLI) version of this program. Links and instructions for doing so are available [here](https://ffmpeg.org/download.html). Below are instructions for several popular platforms.

**Debian/Ubuntu**

1. Enter this code into your terminal:<br />`sudo apt-get install ffmpeg`
2. Check that R can find it:<br />
`tidymedia::program_status()`<br />
The `ffmpeg` and `ffprobe` rows should show locations and versions. If either
is `NA`, tidymedia cannot see that program: give it the path yourself with
`tidymedia::set_ffmpeg("/path/to/ffmpeg")` or
`tidymedia::set_ffprobe("/path/to/ffprobe")`. The two are looked up separately,
so a row that is still `NA` needs its own call.

**Windows**

1. Install the `tidymedia` package in R.
2. Run `tidymedia::install_on_win()` in R and confirm the prompt, which names the digest and the archive it will download, the directory it will unpack into, and the program locations it may overwrite. The archive is checked against the SHA-256 digest gyan.dev publishes beside it before anything is unpacked, and no program location is remembered unless the extraction produced that program. To install from a different build, pass its address and its digest: `install_on_win(download_url = ..., archive_checksum = ...)`.
3. Check that R can find it:<br />
`tidymedia::program_status()`<br />
The `ffmpeg` and `ffprobe` rows should show locations and versions. If either
is `NA`, tidymedia cannot see that program: read the message
`install_on_win()` printed, or point tidymedia at a build you already have with
`tidymedia::set_ffmpeg("C:/path/to/ffmpeg.exe")` or
`tidymedia::set_ffprobe("C:/path/to/ffprobe.exe")`. The two are looked up
separately, so a row that is still `NA` needs its own call.

**macOS Homebrew Install**

1. Open the macOS Terminal or Linux shell prompt.
2. Install [Homebrew](https://brew.sh/) by entering this code into your terminal:<br />`/bin/bash -c "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/HEAD/install.sh)"`
3. Enter this code into your terminal:<br />`brew install ffmpeg`
4. Check that R can find it:<br />
`tidymedia::program_status()`<br />
The `ffmpeg` and `ffprobe` rows should show locations and versions. If either
is `NA`, tidymedia cannot see that program: run `brew --prefix ffmpeg` in the
terminal and give tidymedia the `bin/ffmpeg` and `bin/ffprobe` under what it
prints, with `tidymedia::set_ffmpeg("/path/to/ffmpeg")` and
`tidymedia::set_ffprobe("/path/to/ffprobe")`. The two are looked up separately,
so a row that is still `NA` needs its own call.

**macOS Manual Install**

1. Download the latest snapshot version from:<br /><https://evermeet.cx/ffmpeg/get><br />
That address gives you `ffmpeg` on its own. The metadata functions also need
`ffprobe`, which is a separate download:<br />
<https://evermeet.cx/ffmpeg/getrelease/ffprobe/7z>
2. Extract the contents of each downloaded [.7z](https://www.7-zip.org/) file
3. Drag the extracted contents to the Applications folder
4. The Applications folder is not on the `PATH`, so nothing looking a program up
   by name will find it there. Tell tidymedia where each program landed
   (changing the paths to match Step 3):<br />
`tidymedia::set_ffmpeg("/Applications/ffmpeg")`<br />
`tidymedia::set_ffprobe("/Applications/ffprobe")`<br />
Each asks you to confirm the path before it remembers it, and it is remembered
across R sessions. Set both: the two programs are looked up separately, so
setting `ffmpeg` does not tell tidymedia where `ffprobe` is.
5. Check that R can find them:<br />
`tidymedia::program_status()`<br />
The `ffmpeg` and `ffprobe` rows should show locations and versions. If either is
`NA`, the path you gave in Step 4 is not where that program actually landed:
find it in the Applications folder and run the call again with that path.

## Examples

```{r}
library(tidymedia)
```

The examples below use a tiny sample clip that ships with the package:

```{r}
video <- system.file("extdata", "sample.mp4", package = "tidymedia")
```

### Build reproducible FFmpeg commands

The `ffm_*` builder assembles a command step by step. Nothing runs until you ask it to: `ffm_compile()` returns the exact FFmpeg command as a string, and `ffm_run()` executes it. See `vignette("tidymedia")` for the full tour.

```{r}
ffm_files(video, "output.mp4") |>
  ffm_trim(start = 1, end = 5) |>
  ffm_crop(width = 160, height = 120) |>
  ffm_codec(video = "libx264") |>
  ffm_drop(streams = "audio") |>
  ffm_compile()
```

Common tasks have their own verbs (Layer 2). Pass `run = FALSE` to see the command without executing it:

```{r}
extract_audio(video, "audio.aac", run = FALSE)
```

### Read metadata as tibbles

`probe_all()` returns container- and stream-level metadata from FFprobe as tibbles, keyed by a leading `file` column so a whole batch stacks cleanly:

```{r, eval = has_ffprobe}
probe_all(video)$streams
```

MediaInfo is available too, via `mediainfo_query()`, `mediainfo_template()`, and the `get_*()` shortcuts (see `vignette("metadata")`):

```{r, eval = has_mediainfo}
get_duration(video, unit = "sec")
get_width(video)
```

### Query FFmpeg's capabilities

```{r, eval = has_ffmpeg}
head(ffmpeg_codecs())
```

## Code of Conduct

Please note that the **tidymedia** project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
