---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidymedia

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R-CMD-check](https://github.com/jmgirard/tidymedia/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jmgirard/tidymedia/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/jmgirard/tidymedia/graph/badge.svg)](https://app.codecov.io/gh/jmgirard/tidymedia)
<!-- badges: end -->

The goal of **tidymedia** is to provide tools for easily working with media (e.g., image, audio, and video) files within R and the tidyverse. It wraps [FFmpeg](https://ffmpeg.org/) and [MediaInfo](https://mediaarea.net/en/MediaInfo) for **reproducible media preprocessing** — batch trimming, cropping, format standardization, and metadata extraction as tibbles.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jmgirard/tidymedia")
```

### Dependencies

#### 1. MediaInfo

**tidymedia** uses [MediaInfo](https://mediaarea.net/en/MediaInfo) to query information about media files. If you would like to use these functions, you will need to install the command line interface (CLI) version of this program. Links and instructions for doing so are available [here](https://mediaarea.net/en/MediaInfo/Download). Below are instructions for several popular platforms.

**Debian/Ubuntu**

1. Enter this code into your terminal:<br />`sudo apt-get install mediainfo`

**Windows**

1. Download the appropriate CLI .zip file from:<br /><https://mediaarea.net/en/MediaInfo/Download/Windows>
2. Extract (or copy) the contents of this .zip file to a folder on your computer such as:<br />`C:/Program Files/MediaInfo`
3. Run the following code in R (changing the path to match Step 2):<br />
`tidymedia::set_mediainfo("C:/Program Files/MediaInfo/mediainfo.exe")`

**Mac**

1. Download the appropriate CLI .dmg file from:<br /><https://mediaarea.net/en/MediaInfo/Download/Mac_OS>
2. Open the .dmg file and drag the program icon to the Applications folder

#### 2. FFmpeg
**tidymedia** uses [FFmpeg](https://ffmpeg.org/) to encode media files. If you would like to use these functions, you will need to install the command line interface (CLI) version of this program. Links and instructions for doing so are available [here](https://ffmpeg.org/download.html). Below are instructions for several popular platforms.

**Debian/Ubuntu**

1. Enter this code into your terminal:<br />`sudo apt-get install ffmpeg`

**Windows**

1. Install the `tidymedia` package in R.
2. Run `tidymedia::install_on_win()` in R.

**macOS Homebrew Install**

1. Open the macOS Terminal or Linux shell prompt.
2. Install [Homebrew](https://brew.sh/) by entering this code into your terminal:<br />`/bin/bash -c "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/master/install.sh)"`
3. Enter this code into your terminal:<br />`brew install ffmpeg`

**macOS Manual Install**

1. Download the latest snapshot version from:<br /><https://evermeet.cx/ffmpeg/get>
2. Extract the contents of the downloaded [.7z](https://www.7-zip.org/) file
3. Drag the extracted contents to the Applications folder

## Examples

```{r}
library(tidymedia)
```

The examples below use a tiny sample clip that ships with the package:

```{r}
video <- system.file("extdata", "sample.mp4", package = "tidymedia")
```

### Build reproducible FFmpeg commands

The `ffm_*` builder assembles a command step by step. Nothing runs until you ask it to: `ffm_compile()` returns the exact FFmpeg command as a string, and `ffm_run()` executes it. See `vignette("tidymedia")` for the full tour.

```{r}
ffm(video, "output.mp4") |>
  ffm_trim(start = 1, end = 5) |>
  ffm_crop(width = 160, height = 120) |>
  ffm_codec(video = "libx264") |>
  ffm_drop(streams = "audio") |>
  ffm_compile()
```

Common tasks have their own verbs (Layer 2). Pass `run = FALSE` to see the command without executing it:

```{r}
extract_audio(video, "audio.aac", run = FALSE)
```

### Read metadata as tibbles

`probe_all()` returns container- and stream-level metadata from FFprobe as tibbles, keyed by a leading `file` column so a whole batch stacks cleanly:

```{r}
probe_all(video)$streams
```

MediaInfo is available too, via `mediainfo_query()`, `mediainfo_template()`, and the `get_*()` shortcuts (see `vignette("metadata")`):

```{r}
get_duration(video, unit = "sec")
get_width(video)
```

### Query FFmpeg's capabilities

```{r}
head(ffmpeg_codecs())
```

## Code of Conduct

Please note that the **tidymedia** project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
