---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
# Some of the chunks below run real commands, which means starting FFmpeg,
# FFprobe or MediaInfo. Each of those is evaluated only when the program it
# needs is installed, so this README knits on a machine that has none of them --
# the machine every reader of the install section above is on. Apart from this
# one, five chunks carry no guard, because they start nothing: loading the package, copying the
# sample clip, and the three that only COMPILE commands (the builder, the
# `extract_audio()` call, and the batch example). A reader without the binaries
# should still see what those produce.
# `tools/vignette_chunk_guards.R` is what says which chunk is which.
#
# The examples run in a scratch folder that the sample clip is copied into, so
# what they print names `sample.mp4` rather than the package's install path --
# which `devtools::build_readme()` changes on every run, and which would ship a
# local path in README.md.
readme_dir <- tempfile("tidymedia-readme-")
dir.create(readme_dir)
knitr::opts_knit$set(root.dir = readme_dir)
has_ffmpeg <- nzchar(Sys.which("ffmpeg"))
has_ffprobe <- nzchar(Sys.which("ffprobe"))
has_mediainfo <- nzchar(Sys.which("mediainfo"))
# Recorded on stderr rather than into the page: this line is what a build log
# shows for the three programs, so a build that was meant to run without them
# can be checked rather than assumed. `message()` would be captured by knitr and
# then dropped with the rest of this chunk's output.
writeLines(
  paste0("tidymedia README build, Sys.which(): ", paste0(
    c("ffmpeg", "ffprobe", "mediainfo"), "=[",
    Sys.which(c("ffmpeg", "ffprobe", "mediainfo")), "]", collapse = " ")),
  con = stderr()
)
```

# tidymedia

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/jmgirard/tidymedia/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jmgirard/tidymedia/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/jmgirard/tidymedia/graph/badge.svg)](https://app.codecov.io/gh/jmgirard/tidymedia)
<!-- badges: end -->

**tidymedia** helps you work with video, audio and image files in R. It runs
[FFmpeg](https://ffmpeg.org/) and [MediaInfo](https://mediaarea.net/en/MediaInfo)
for you, so you can prepare media for research in a way you can repeat. It
trims, crops and converts files, often many at once. It also reads media
metadata into tibbles.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jmgirard/tidymedia")
```

### Dependencies

tidymedia uses two free command-line programs. [FFmpeg](https://ffmpeg.org/)
converts media files and comes with FFprobe, which reads them.
[MediaInfo](https://mediaarea.net/en/MediaInfo) also reads media files. Install
the ones you need.

**Debian and Ubuntu.** In a terminal, run:<br />
`sudo apt-get install ffmpeg mediainfo`

**macOS.** Install [Homebrew](https://brew.sh/). Then, in a terminal, run:<br />
`brew install ffmpeg media-info`

**Windows.** For FFmpeg, run `tidymedia::install_on_win()` in R. It asks you
to confirm before it downloads anything. For MediaInfo, download the CLI
version from the [MediaInfo download page](https://mediaarea.net/en/MediaInfo/Download/Windows)
and unzip it. Then tell tidymedia where the program is:<br />
`tidymedia::set_mediainfo("C:/Program Files/MediaInfo/mediainfo.exe")`

**Check the install.** In R, run:<br />
`tidymedia::program_status()`

Each program that tidymedia found shows a location and a version. If a location
is `NA`, give tidymedia the path with `set_ffmpeg()`, `set_ffprobe()` or
`set_mediainfo()`. The help pages `?set_program` and `?install_on_win` give the
details.

## Examples

```{r}
library(tidymedia)
```

The examples below use a short sample clip that comes with the package. They
copy it to the working folder, so the paths they print stay short:

```{r}
invisible(file.copy(system.file("extdata", "sample.mp4", package = "tidymedia"), "."))
video <- "sample.mp4"
```

### Build reproducible FFmpeg commands

The pipeline functions, whose names start with `ffm_`, build an FFmpeg command
one step at a time. Nothing runs until you ask. `ffm_compile()` returns the
command as a string, and `ffm_run()` runs it. See `vignette("tidymedia")` for a
full tour.

```{r}
ffm_files(video, "output.mp4") |>
  ffm_trim(start = 1, end = 5) |>
  ffm_crop(width = 160, height = 120) |>
  ffm_codec(video = "libx264") |>
  ffm_drop(streams = "audio") |>
  ffm_compile()
```

Common jobs have their own task functions. Add `run = FALSE` to see the command
without running it:

```{r}
extract_audio(video, "audio.aac", run = FALSE)
```

### Process a folder in batch

`ffm_jobs()` turns a folder into a jobs table, with one row for each media file.
Some `*_batch()` functions, such as `crop_video_batch()`, take that table as it
is. Others need a column added first, such as `output`.

If the folder has no files of the type you ask for, `ffm_jobs()` stops with an
error. See `vignette("batch")` for more.

```{r}
jobs <- ffm_jobs(".", type = "video")
# ffm_jobs() returns full paths; keep only the file name so this page stays
# short. The clip is in the working folder, so the commands still find it.
jobs$input <- basename(jobs$input)
crop_video_batch(jobs, width = 160, height = 120, run = FALSE)
```

### Read metadata as tibbles

`probe_all()` uses FFprobe to read facts about the
[container](https://jmgirard.github.io/tidymedia/articles/tidymedia.html#glossary)
and each [stream](https://jmgirard.github.io/tidymedia/articles/tidymedia.html#glossary).
It returns them as tibbles. Each tibble starts with a `file` column, so the
results for many files stack into one table:

```{r, eval = has_ffprobe}
probe_all(video)$streams
```

MediaInfo works too, through `mediainfo_query()`, `mediainfo_template()` and
the `get_*()` functions. See `vignette("metadata")`.

```{r, eval = has_mediainfo}
get_duration(video, unit = "sec")
get_width(video)
```

### Query FFmpeg's capabilities

`ffmpeg_codecs()` lists the
[codecs](https://jmgirard.github.io/tidymedia/articles/tidymedia.html#glossary)
that your FFmpeg build knows:

```{r, eval = has_ffmpeg}
head(ffmpeg_codecs())
```

## Code of Conduct

Please note that the **tidymedia** project is released with a [Contributor Code of Conduct](https://www.contributor-covenant.org/version/2/0/code_of_conduct/). By contributing to this project, you agree to abide by its terms.
