---
title: "Batch processing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Batch processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidymedia)
```

Batch processing over many files is what tidymedia is built for. The examples
below use `run = FALSE`, which **compiles** each FFmpeg command without
executing it — so you can inspect exactly what would run. Drop that argument (or
set `run = TRUE`) to actually process the files.

```{r}
folder <- system.file("extdata", package = "tidymedia")
video <- system.file("extdata", "sample.mp4", package = "tidymedia")
```

## The batch runner

`ffm_batch()` is the general-purpose engine: it takes a **jobs tibble** (one row
per output) and a function that turns a row into a pipeline. Each column is
passed to the function by name, `purrr::pmap()`-style, and the runner returns
your jobs tibble with a `command` column added (and a `success` column when
`run = TRUE`).

`ffm_jobs()` builds that tibble from a folder: one row per file of the media
type you ask for, with the file's full path in an `input` column. When the
folder holds no file of that type, `ffm_jobs()` stops with an error rather than
returning an empty table. Add any other columns your function needs, here an
`output` for each input:

```{r}
jobs <- ffm_jobs(folder, type = "video")
jobs$output <- paste0(tools::file_path_sans_ext(basename(jobs$input)), ".mp3")

ffm_batch(jobs, run = FALSE, .f = function(input, output, ...) {
  ffm_files(input, output) |>
    ffm_drop("video") |>
    ffm_codec(audio = "libmp3lame")
})
```

Because you build the pipeline yourself, any combination of builder verbs is
available per job. Give `.f` a `...` argument so it tolerates extra job columns
it does not use: without one, a column `.f` has no argument for stops the batch
with R's "unused argument" error.

## Per-verb batch siblings

You do not have to write `.f` by hand for the common jobs. Every task verb ships
a `*_batch()` companion — `extract_audio_batch()`, `convert_audio_batch()`,
`crop_video_batch()`, `standardize_video_batch()`, `normalize_audio_batch()`,
and the rest — that takes a jobs tibble directly and applies the verb to each
row. `crop_video_batch()` can take the table straight from `ffm_jobs()`;
others, such as `extract_audio_batch()`, refuse it until you add an `output`
column. `ffm_jobs()` again stops with an error rather than returning an empty
table when the folder holds no file of the type:

```{r}
jobs <- ffm_jobs(folder, type = "video")
jobs$output <- paste0(tools::file_path_sans_ext(basename(jobs$input)), "_cropped.mp4")

crop_video_batch(jobs, width = 160, height = 120, run = FALSE)
```

Without an `output` column, `crop_video_batch()` names each output after its
input, adding `_cropped`, in the input's own folder. Here that folder is inside
the installed package, so the example adds an `output` column that puts the
files in the working directory instead. Either way `crop_video_batch()` rejects
two rows that would resolve to the same output path. `vignette("workflow")` chains several of
these across a study folder.

## Fan-out verbs

Some tasks turn one input into *many* outputs. These fan-out verbs are Layer 2
wrappers built on `ffm_batch()`.

`segment_video()` cuts a file into pieces given start/stop timestamps, returning
one row per segment:

```{r}
segment_video(
  video,
  start = c(0, 0.5),
  end  = c(0.5, 1),
  run = FALSE
)
```

`separate_audio_video()` splits a file into its audio and video streams,
returning the two compiled commands:

```{r}
separate_audio_video(video, "audio.aac", "video.mp4", run = FALSE)
```

## Running in parallel

`parallel = TRUE` is accepted by `ffm_batch()` and by every `*_batch` verb; by
`segment_video()`, which sends its own segments through the same machinery; and
by the five metadata readers `probe_all()`, `probe_container()`,
`probe_streams()`, `probe_video()`, and `probe_audio()`. On the four reader
shortcuts it applies when you pass `infile` and is ignored when you hand them an
existing `probe` object, which has nothing left to fan out. Nothing else takes
the argument — `separate_audio_video()` compiles its two commands directly
rather than through `ffm_batch()`, so only `separate_audio_video_batch()`
accepts it.

The argument maps over jobs with [furrr](https://furrr.futureverse.org/), and
parallelism follows whatever [future](https://future.futureverse.org/) plan you
have set, so you opt in explicitly. With no plan set, `parallel = TRUE` still
runs one job at a time, and warns to tell you so:

```{r, eval = FALSE}
library(future)
plan(multisession)

ffm_batch(jobs, parallel = TRUE, .f = function(input, output, ...) {
  ffm_files(input, output) |> ffm_drop("video") |> ffm_codec(audio = "libmp3lame")
})
```

Every job carries its own compiled command in the returned tibble, so a batch
run is fully reproducible: save that column and you have an exact record of the
FFmpeg commands that produced your outputs.

## Where to next

- `vignette("workflow")` — an end-to-end research pipeline that applies these
  batch tools across a study folder.
- `vignette("tidymedia")` — the task verbs and the builder they are made of.
- `vignette("metadata")` — reading each file's metadata as a tibble.
