---
title: "Batch processing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Batch processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidymedia)
```

Batch processing over many files is what tidymedia is built for. The examples
below use `run = FALSE`, which **compiles** each FFmpeg command without
executing it — so you can inspect exactly what would run. Drop that argument (or
set `run = TRUE`) to actually process the files.

```{r}
video <- system.file("extdata", "sample.mp4", package = "tidymedia")
```

## The batch runner

`ffm_batch()` is the general-purpose engine: it takes a **jobs tibble** (one row
per output) and a function that turns a row into a pipeline. Each column is
passed to the function by name, `purrr::pmap()`-style, and the runner returns
your jobs tibble with a `command` column added (and a `success` column when
`run = TRUE`).

```{r}
jobs <- tibble::tibble(
  input  = c(video, video),
  output = c("clip1.mp3", "clip2.mp3")
)

ffm_batch(jobs, run = FALSE, .f = function(input, output, ...) {
  ffm_files(input, output) |>
    ffm_drop("video") |>
    ffm_codec(audio = "libmp3lame")
})
```

Because you build the pipeline yourself, any combination of builder verbs is
available per job. Give `.f` a `...` argument so it tolerates extra job columns
it does not use.

## Per-verb batch siblings

You do not have to write `.f` by hand for the common jobs. Every task verb ships
a `*_batch()` companion — `extract_audio_batch()`, `convert_audio_batch()`,
`crop_video_batch()`, `standardize_video_batch()`, `normalize_audio_batch()`,
and the rest — that takes a jobs tibble directly and applies the verb to each
row:

```{r}
jobs <- tibble::tibble(
  input  = c(video, video),
  output = c("session01_cropped.mp4", "session02_cropped.mp4")
)

crop_video_batch(jobs, width = 160, height = 120, run = FALSE)
```

When the jobs tibble has no `output` column, the transform verbs auto-name each
output from its input (e.g. `_cropped`); either way they reject two rows that
would resolve to the same output path. `vignette("workflow")` chains several of
these across a study folder.

## Fan-out verbs

Some tasks turn one input into *many* outputs. These fan-out verbs are Layer 2
wrappers built on `ffm_batch()`.

`segment_video()` cuts a file into pieces given start/stop timestamps, returning
one row per segment:

```{r}
segment_video(
  video,
  start = c(0, 0.5),
  end  = c(0.5, 1),
  run = FALSE
)
```

`separate_audio_video()` splits a file into its audio and video streams,
returning the two compiled commands:

```{r}
separate_audio_video(video, "audio.aac", "video.mp4", run = FALSE)
```

## Running in parallel

Both `ffm_batch()` and the fan-out verbs accept `parallel = TRUE`, which maps
over jobs with [furrr](https://furrr.futureverse.org/). Parallelism follows
whatever [future](https://future.futureverse.org/) plan you have set, so you opt
in explicitly:

```{r, eval = FALSE}
library(future)
plan(multisession)

ffm_batch(jobs, parallel = TRUE, .f = function(input, output, ...) {
  ffm_files(input, output) |> ffm_drop("video") |> ffm_codec(audio = "libmp3lame")
})
```

Every job carries its own compiled command in the returned tibble, so a batch
run is fully reproducible: save that column and you have an exact record of the
FFmpeg commands that produced your outputs.

## Where to next

- `vignette("workflow")` — an end-to-end research pipeline that applies these
  batch tools across a study folder.
- `vignette("tidymedia")` — the task verbs and the builder they are made of.
- `vignette("metadata")` — reading each file's metadata as a tibble.
