---
title: "Media metadata as tibbles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Media metadata as tibbles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# Metadata readers shell out to FFprobe / MediaInfo. Chunks that need a binary
# are evaluated only when it is available, so this vignette builds cleanly on
# machines (and CI images) that lack the command-line tools.
has_ffprobe <- nzchar(Sys.which("ffprobe"))
has_mediainfo <- nzchar(Sys.which("mediainfo"))
```

```{r setup}
library(tidymedia)
```

tidymedia reads media metadata as **tibbles**, so a whole folder of files
becomes a data frame you can filter, join, and summarize with the tidyverse.
Two engines back these readers:

- **FFprobe** (`probe_*()`) returns container- and stream-level metadata.
- **MediaInfo** (`mediainfo_*()`, `get_*()`) returns a broader, differently
  organized set of fields.

Both require the corresponding command-line tool to be installed (see the
[README](https://github.com/jmgirard/tidymedia) for setup). We use the sample
clip that ships with the package:

```{r}
video <- system.file("extdata", "sample.mp4", package = "tidymedia")
```

## Probing with FFprobe

`probe_all()` returns a list of two tibbles: `container` (one row per file) and
`streams` (one row per stream). Both lead with a `file` column, so results
stack cleanly when you pass several files.

```{r, eval = has_ffprobe}
info <- probe_all(video)
info$container
```

```{r, eval = has_ffprobe}
info$streams
```

The `probe_*()` shortcuts pull out just what you need. They accept **either** a
`probe` object (to avoid re-reading the file) or a file location via `infile`:

```{r, eval = has_ffprobe}
# Reuse the probe object rather than reprobing
probe_video(info)
```

Numeric columns are typed by default (`typed = TRUE`); pass `typed = FALSE` to
keep everything as strings. Frame rates that FFprobe reports as fractions (e.g.
`"30000/1001"`) are evaluated to doubles automatically when the columns are
typed.

## Querying with MediaInfo

MediaInfo organizes metadata into sections (`General`, `Video`, `Audio`, ...).
`mediainfo_query()` pulls several parameters from one section into a tibble:

```{r, eval = has_mediainfo}
mediainfo_query(
  video,
  section = "Video",
  parameters = c("Width", "Height", "FrameRate")
)
```

`mediainfo_template()` applies a whole template at once; two ship with the
package (`"brief"` and `"extended"`):

```{r, eval = has_mediainfo}
mediainfo_template(video, template = "brief")
```

For single values, the `get_*()` helpers are the quickest path:

```{r, eval = has_mediainfo}
get_duration(video, unit = "sec")
get_width(video)
get_height(video)
```

## Batching over many files

Because every reader accepts a vector of files and keys its output by `file`,
describing a whole directory is a one-liner:

```{r, eval = FALSE}
files <- list.files("my/videos", pattern = "\\.mp4$", full.names = TRUE)
probe_all(files)$container
```

Files that cannot be read yield an all-`NA` row plus a warning, rather than
aborting the whole call — so one bad file never sinks a batch.
