---
title: "Get started with tidymedia"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with tidymedia}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidymedia)
```

tidymedia helps you build **reproducible media-preprocessing pipelines** on top
of [FFmpeg](https://ffmpeg.org/) and [MediaInfo](https://mediaarea.net/en/MediaInfo).
It is deliberately *not* "all of FFmpeg in R": the focus is batch trimming,
cropping, format standardization, and metadata extraction for research and
data-science work.

Throughout this vignette we use a tiny sample clip that ships with the package:

```{r}
video <- system.file("extdata", "sample.mp4", package = "tidymedia")
```

## Start with a task verb

Most preprocessing jobs are a single call to a **task verb**. Say you need the
audio track of a recording — to hand to a transcription tool, for example.
`extract_audio()` does exactly that:

```{r, eval = nzchar(Sys.which("ffmpeg"))}
extract_audio(video, "audio.m4a")
```

Every task verb runs FFmpeg immediately and returns the path it wrote. To see
the command *without* running anything, pass `run = FALSE`; the verb compiles
the exact FFmpeg invocation and returns it as a string you can inspect, log, or
save:

```{r}
extract_audio(video, "audio.m4a", run = FALSE)
```

That reproducible command is the thread running through the whole package: a
pipeline is a value you can examine before you commit to it. Cropping works the
same way — describe the job, inspect the command, then run it:

```{r}
crop_video(video, "cropped.mp4", width = 160, height = 120, run = FALSE)
```

For preprocessing a whole folder of files at once, every task verb has a batch
sibling (`extract_audio_batch()`, `crop_video_batch()`, …); see
`vignette("batch")`. For an end-to-end research pipeline that chains many verbs
together, see `vignette("workflow")`.

## The three layers

Task verbs are the top of three layers, from highest level to lowest:

- **Layer 2 — the task verbs.** Functions like `extract_audio()` and
  `crop_video()` are thin wrappers for common jobs. This is where most work
  happens; you may never need to look below it.
- **Layer 1 — the pipeline builder.** The `ffm_*` functions assemble an FFmpeg
  command step by step. Every task verb is built from these, and all of the
  quoting, option ordering, and copy-vs-re-encode logic lives here. Reach for
  the builder when no single task verb fits.
- **Layer 0 — the escape hatch.** `ffmpeg()`, `ffprobe()`, and `mediainfo()`
  pass a raw argument string straight to the corresponding command-line tool,
  for anything tidymedia does not wrap.

The rest of this vignette drops down to Layer 1, the builder that every task
verb is made of.

## Building a pipeline

Every pipeline starts with `ffm()` (an alias of `ffm_files()`), which names the
input and output files. You then add steps with `|>`. Each `ffm_*` verb records
an instruction; nothing runs until you ask it to. Calling `ffm_compile()` turns
the pipeline into the exact FFmpeg command it represents:

```{r}
ffm(video, "output.mp4") |>
  ffm_trim(start = 1, end = 5) |>
  ffm_crop(width = 160, height = 120) |>
  ffm_codec(video = "libx264") |>
  ffm_compile()
```

Printing a pipeline shows the same compiled command, so you can eyeball a
pipeline at any point:

```{r}
ffm(video, "output.mp4") |>
  ffm_scale(width = 320, height = 240) |>
  ffm_pixel_format("yuv420p")
```

When you are ready, swap `ffm_compile()` for `ffm_run()` to execute the command
through FFmpeg and write the output file.

## Copy versus re-encode

Cutting a clip can be **frame-accurate** (re-encoded, slower) or **fast and
lossless** (a stream copy that snaps to keyframes). `ffm_seek()` exposes both
via its `reencode` argument, and pairs naturally with `ffm_copy()` for the fast
path:

```{r}
# Fast, lossless cut
ffm(video, "output.mp4") |>
  ffm_seek(start = 1, end = 5, reencode = FALSE) |>
  ffm_copy() |>
  ffm_compile()
```

Note that `ffm_seek()` (which uses FFmpeg's `-ss`/`-to` options) is distinct
from `ffm_trim()` (the `trim` *filter*): only seeking can stream-copy.

## Combining multiple inputs

A few "blessed" verbs take more than one input. Pass a vector of files to
`ffm()`, then combine them: stack side by side with `ffm_hstack()` or top to
bottom with `ffm_vstack()`, composite one over another with `ffm_overlay()`, or
join them end-to-end with `ffm_concat()`:

```{r}
ffm(c(video, video), "side_by_side.mp4") |>
  ffm_hstack() |>
  ffm_compile()
```

Each manages its own stream labels internally, so audio is dropped unless you
map it back with `ffm_map("0:a")`. The task-verb layer wraps the common cases:
`compare_videos()` for a side-by-side (or stacked) comparison and
`picture_in_picture()` for an inset overlay.

```{r}
compare_videos(c(video, video), "compare.mp4", audio = 0, run = FALSE)
```

The builder stays deliberately linear — one input chain, sequential filters,
one output — so full FFmpeg filtergraphs are out of scope. When you need one,
reach for the Layer 0 escape hatch:

```{r, eval = nzchar(Sys.which("ffmpeg"))}
# Layer 0: raw arguments passed straight to FFmpeg
ffmpeg("-version")[1]
```

## Where to next

- `vignette("workflow")` — an end-to-end research preprocessing pipeline.
- `vignette("batch")` — running a verb over many files at once.
- `vignette("metadata")` — reading container and stream metadata as tibbles.
