---
title: "Checking results and bounding runs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Checking results and bounding runs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# The chunks below check and produce real files, which means starting FFmpeg or
# FFprobe. Each is evaluated only when the program it needs is installed, so
# this vignette builds on machines and CI images that have neither.
has_ffmpeg <- nzchar(Sys.which("ffmpeg"))
has_ffprobe <- nzchar(Sys.which("ffprobe"))
has_both <- has_ffmpeg && has_ffprobe
# Recorded on stderr rather than into the page: this line is what a build log
# shows for the three programs, so a build that was meant to run without them
# can be checked rather than assumed. `message()` would be captured by knitr and
# then dropped with the rest of this chunk's output.
writeLines(
  paste0("tidymedia vignette build, Sys.which(): ", paste0(
    c("ffmpeg", "ffprobe", "mediainfo"), "=[",
    Sys.which(c("ffmpeg", "ffprobe", "mediainfo")), "]", collapse = " ")),
  con = stderr()
)
# Work in a temporary directory: nothing is written beside the vignette sources,
# and the paths in the reports below stay short enough to read.
knitr::opts_knit$set(root.dir = tempdir())
```

```{r setup}
library(tidymedia)
```

This page answers three questions about processed files. Does an output have
the properties you asked for? Can you later say how a file was made? And how
do you stop a program that hangs?

The examples use a copy of the sample clip that comes with the package:

```{r}
file.copy(
  system.file("extdata", "sample.mp4", package = "tidymedia"),
  "session01.mp4"
)
```

## Check the output against what you asked for

A task function returns the FFmpeg command it ran. That command says what you
asked for. `verify_media()` reads the output file and says what you got:

```{r, eval = has_both}
extract_audio("session01.mp4", "session01.m4a")

verify_media("session01.m4a", audio_codec = "aac", duration = 1)
```

The result is a tibble with one row for each check. Its columns are `file`,
`check`, `expected`, `actual` and `pass`. When a check fails, `actual` shows
what the file holds instead:

```{r, eval = has_ffprobe}
verify_media("session01.mp4", duration = 5, width = 1920, audio_codec = "aac")
```

You can check any field that FFprobe reports, by its name:

```{r, eval = has_ffprobe}
verify_media("session01.mp4", pix_fmt = "yuv420p", nb_streams = 2)
```

### What the checks cover

The checks read the file's metadata, such as its duration, size and
[codec](tidymedia.html#glossary) names. They do not judge whether the picture
or the sound is good.

A number passes when it is within `tolerance` of the expected value. The
default `tolerance` is `0.1`. So a width or a
[sample rate](tidymedia.html#glossary) must match exactly, but a duration can
differ a little. A codec name must match exactly.

A property that the file does not have fails. For example, the audio file above
has no video [stream](tidymedia.html#glossary), so a width check fails with an
`actual` of `NA`:

```{r, eval = has_both}
verify_media("session01.m4a", width = 320, audio_codec = "aac")
```

Another field name is looked for in the [container](tidymedia.html#glossary)
information first. Then it is looked for in the video stream, then in the audio
stream. The first match is used.

### Checking every job in a batch

`ffm_batch()` and the `*_batch()` task functions take a `verify` argument. It
adds a `verified` column to the result. The column is `TRUE` or `FALSE` for
each job that ran and could be checked. It is `NA` for a job that failed or
could not be checked. A named list applies the same checks to every job:

```{r, eval = has_both}
jobs <- tibble::tibble(
  input  = "session01.mp4",
  output = "session01.mp3"
)

ffm_batch(
  jobs,
  verify = list(audio_codec = "mp3"),
  .f = function(input, output, ...) {
    ffm_files(input, output) |>
      ffm_drop("video") |>
      ffm_codec(audio = "libmp3lame")
  }
)
```

A failed check does not stop the batch. The other jobs still run, and the
`verified` column shows which outputs to look at. To use different checks for
each job, pass a function of the job columns in place of the list.

## Record how the files were made

The command for each job records what you asked FFmpeg to do. It does not
record which FFmpeg version ran, and two versions can make different files from
one command.

`ffm_batch(manifest = TRUE)` records those facts. The manifest is stored with
the batch result, and `ffm_manifest()` reads it:

```{r, eval = has_both}
res <- ffm_batch(
  jobs,
  manifest = TRUE,
  .f = function(input, output, ...) {
    ffm_files(input, output) |>
      ffm_drop("video") |>
      ffm_codec(audio = "libmp3lame")
  }
)

res

ffm_manifest(res)
```

The manifest has one row for each job. The `command`, `input` and `output`
columns repeat the job. The other columns record the run:

- `ffmpeg_version` is the FFmpeg version that did the work, and
  `ffprobe_version` is the FFprobe version that was available. tidymedia reads
  both from the programs. A version it cannot read is `NA`.
- `timestamp` is the time of the run, with its offset from UTC.
- `output_size` is the size of the output file in bytes, so an empty or cut-off
  file shows in the record.

A job with several inputs lists them in one `input` cell, separated by `;`.

With `checksums = TRUE`, each row also has `input_md5` and `output_md5`. A
checksum lets you show later that a file on disk is the file in the record:

```{r, eval = has_both}
res <- ffm_batch(
  jobs,
  manifest = TRUE,
  checksums = TRUE,
  .f = function(input, output, ...) {
    ffm_files(input, output) |>
      ffm_drop("video") |>
      ffm_codec(audio = "libmp3lame")
  }
)

ffm_manifest(res)[, c("input_md5", "output_md5")]
```

Checksums are off by default, because they read every input and output file in
full. That takes time on a large study.

The `*_batch()` task functions take `manifest` and `checksums` too. So
`extract_audio_batch()` and the others record the same facts.

If you give `ffm_manifest()` a `path`, it also writes the manifest to a CSV
file. You can then keep that file next to the processed data:

```{r, eval = has_both}
ffm_manifest(res, path = "session01_manifest.csv")

file.exists("session01_manifest.csv")
```

A batch run with `run = FALSE` runs nothing, so it has no manifest. Reading the
manifest of such a batch is an error, not an empty tibble:

```{r, eval = has_both, error = TRUE}
compiled <- ffm_batch(jobs, run = FALSE, manifest = TRUE, .f = function(input, output, ...) {
  ffm_files(input, output) |> ffm_drop("video")
})

ffm_manifest(compiled)
```

## Bound a run that hangs

Each tidymedia call waits for the program it started. If that program hangs,
the R session waits with it. For example, a network drive can stall, or FFmpeg
can get stuck on a damaged file. A time limit stops the call instead.

The limit is an option, in whole seconds:

```{r, eval = FALSE}
options(tidymedia.timeout = 600)
```

The default is `0`, which means no limit.

To limit one call, wrap it in `with_timeout()`. When the call ends, the option
goes back to its earlier value:

```{r, eval = has_both}
with_timeout(
  extract_audio("session01.mp4", "bounded.m4a"),
  seconds = 60
)

getOption("tidymedia.timeout")
```

To limit the rest of a function, call `local_timeout()` inside it. Each program
started after that line has the limit. When the function returns, the option
goes back to its earlier value:

```{r, eval = has_both}
convert_all <- function(files) {
  local_timeout(60)
  vapply(files, function(f) extract_audio(f, sub("[.]mp4$", ".m4a", f)),
         character(1))
}

convert_all("session01.mp4")

getOption("tidymedia.timeout")
```

### Limits of the time limit

The limit sets how long R waits, not how long the program runs. A program that
does not stop when asked can make R wait up to 40 seconds longer than the
limit. So choose a limit with room to spare.

The limit applies to each program that tidymedia starts. In a batch of 100
jobs, each job gets the full limit.

The limit must be a whole number of seconds. A value such as `0.5` is an error:

```{r, error = TRUE}
with_timeout(extract_audio("session01.mp4", "x.m4a"), seconds = 0.5)
```

### What happens when the limit is reached

A reached limit always gives an error or a warning. The task functions,
`ffm_run()`, the direct commands and `verify_media()` stop with an error.

Functions that work on many files give a warning, so the other files are not
lost. The metadata readers return a row of `NA` values. `ffm_batch()` and the
`*_batch()` task functions mark the job with `success = FALSE`.

To handle these in code, use the condition classes:

- `tidymedia_timeout` for the error.
- `tidymedia_probe_timeout` for a warning from a version check or a track
  check.
- `tidymedia_batch_timeout` for the warning from a batch.

The warning from a metadata reader has no condition class of its own.

`?tidymedia` lists which function gives which condition.

## Where to next

- `vignette("workflow")` shows a full research example.
- `vignette("batch")` shows how to run a task function over many files.
- `vignette("metadata")` shows how to read metadata into tibbles.
- `?tidymedia` lists the package options in one place.
