---
title: "Checking results and bounding runs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Checking results and bounding runs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# The chunks below check and produce real files, which means starting FFmpeg or
# FFprobe. Each is evaluated only when the program it needs is installed, so
# this vignette builds on machines and CI images that have neither.
has_ffmpeg <- nzchar(Sys.which("ffmpeg"))
has_ffprobe <- nzchar(Sys.which("ffprobe"))
has_both <- has_ffmpeg && has_ffprobe
# Recorded on stderr rather than into the page: this line is what a build log
# shows for the three programs, so a build that was meant to run without them
# can be checked rather than assumed. `message()` would be captured by knitr and
# then dropped with the rest of this chunk's output.
writeLines(
  paste0("tidymedia vignette build, Sys.which(): ", paste0(
    c("ffmpeg", "ffprobe", "mediainfo"), "=[",
    Sys.which(c("ffmpeg", "ffprobe", "mediainfo")), "]", collapse = " ")),
  con = stderr()
)
# Work in a temporary directory: nothing is written beside the vignette sources,
# and the paths in the reports below stay short enough to read.
knitr::opts_knit$set(root.dir = tempdir())
```

```{r setup}
library(tidymedia)
```

A processed file raises two questions that the pipeline itself cannot answer.
Did the encode actually produce what you asked for? And a year later, can you
say how a given file was made? This vignette covers the three facilities that
answer them — checking an output against expectations, recording a run's
provenance, and bounding a run that hangs so it stops the call instead of your
session.

We work on a copy of the sample clip that ships with the package:

```{r}
file.copy(
  system.file("extdata", "sample.mp4", package = "tidymedia"),
  "session01.mp4"
)
```

## Check the output against what you asked for

Every verb returns the FFmpeg command it compiled, and that command is a record
of what you *asked* for. `verify_media()` answers the other half: it probes a
file and compares its actual properties against your expectations, one row per
check.

```{r, eval = has_both}
extract_audio("session01.mp4", "session01.m4a")

verify_media("session01.m4a", audio_codec = "aac", duration = 1)
```

The report is a tibble with `file`, `check`, `expected`, `actual` and `pass`, so
a failure tells you which property was wrong and what the file holds instead:

```{r, eval = has_ffprobe}
verify_media("session01.mp4", duration = 5, width = 1920, audio_codec = "aac")
```

Three things are worth knowing before you rely on it.

**The checks are structural, not perceptual.** They read FFprobe's metadata —
duration, dimensions, codec names, sample rate — and say nothing about whether
the picture or the sound looks and sounds right. An encode that came out at the
right size in the right codec and looks terrible passes every check above.

**Numeric checks carry a tolerance; string checks do not.** A numeric
expectation passes when `abs(actual - expected) <= tolerance`, and `tolerance`
defaults to `0.1`. That is exact for integers like width and sample rate, and a
little slack for `duration`, which moves when a cut snaps to a keyframe. Codec
names must match exactly.

**A property that is not there fails.** If the file has no stream of that kind,
or FFprobe reports no such field, the actual value is `NA` and the check fails
rather than being skipped. The audio file extracted above has no picture, so
asking it for a width fails with nothing to report:

```{r, eval = has_both}
verify_media("session01.m4a", width = 320, audio_codec = "aac")
```

Beyond the named arguments, any FFprobe field can be checked by name. Extra
names are resolved against the container first, then the video stream, then the
audio stream, and the first match wins:

```{r, eval = has_ffprobe}
verify_media("session01.mp4", pix_fmt = "yuv420p", nb_streams = 2)
```

### Checking every job in a batch

`ffm_batch()` and the `*_batch()` verbs take the same expectations through a
`verify =` argument and add a `verified` column to the result, one logical per
job. Pass a named list to hold every job to the same expectations:

```{r, eval = has_both}
jobs <- tibble::tibble(
  input  = "session01.mp4",
  output = "session01.mp3"
)

ffm_batch(
  jobs,
  verify = list(audio_codec = "mp3"),
  .f = function(input, output, ...) {
    ffm_files(input, output) |>
      ffm_drop("video") |>
      ffm_codec(audio = "libmp3lame")
  }
)
```

A failed check here marks the row and does not stop the batch — the remaining
jobs still run, and `verified` tells you afterwards which outputs to look at.
For expectations that differ per job, pass a function of the job columns
instead of a list, and it is called once per row.

## Record how the files were made

A compiled command is a complete recipe, and `vignette("workflow")` builds its
reproducibility story on exactly that: capture the command each verb returns —
or the `command` column the batch runner adds — and you can re-run any step. The
command is what you *asked* for, though, and it is silent about the run itself.
Two FFmpeg versions accept the same command and need not produce the same file,
which is why the manifest below records the version that ran.

`ffm_batch(manifest = TRUE)` records the rest. The result itself looks like any
other batch result — the manifest rides along on it as an attribute, and
`ffm_manifest()` reads it out:

```{r, eval = has_both}
res <- ffm_batch(
  jobs,
  manifest = TRUE,
  .f = function(input, output, ...) {
    ffm_files(input, output) |>
      ffm_drop("video") |>
      ffm_codec(audio = "libmp3lame")
  }
)

res

ffm_manifest(res)
```

One row per job. `command`, `input` and `output` restate the job; the other
columns are what the command cannot carry:

- `ffmpeg_version` and `ffprobe_version` — the versions tidymedia resolved when
  the batch ran, read out of the binaries themselves rather than assumed.
  `ffmpeg_version` is the version that did the work; `ffprobe_version` is the
  one that was on hand to probe, which a job like this one never needed. A
  version that could not be read is `NA`.
- `timestamp` — when the run happened, with its UTC offset.
- `output_size` — the size in bytes of the file that came out, so a truncated or
  empty result is visible in the record itself.

Multiple inputs — a stacked or concatenated job — are joined by `;` in the one
`input` cell.

Ask for `checksums = TRUE` and each row also carries `input_md5` and
`output_md5`, which is what lets you say later that a file on disk is the file
this row describes:

```{r, eval = has_both}
res <- ffm_batch(
  jobs,
  manifest = TRUE,
  checksums = TRUE,
  .f = function(input, output, ...) {
    ffm_files(input, output) |>
      ffm_drop("video") |>
      ffm_codec(audio = "libmp3lame")
  }
)

ffm_manifest(res)

ffm_manifest(res)[, c("input_md5", "output_md5")]
```

The manifest is nine columns wide now, so at a typical console width the two new
ones print as a footer line rather than in the table; the second call above pulls
them out to show what they hold.

Checksums are off by default because they read every input and output in full,
which is real time on a large study.

`manifest =` and `checksums =` reach the `*_batch()` verbs too, so a pipeline
built from `extract_audio_batch()` and its siblings records provenance the same
way without dropping down to `ffm_batch()`.

Give `ffm_manifest()` a `path` and it writes the manifest as CSV as well as
returning it, invisibly — one file to commit next to the processed data:

```{r, eval = has_both}
ffm_manifest(res, path = "session01_manifest.csv")

file.exists("session01_manifest.csv")
```

A manifest is a record of a run, so there is nothing to record when nothing ran.
A `run = FALSE` batch attaches none even when you ask for one, and reading it
back is an error rather than an empty tibble — a silent empty record being the
failure a provenance record exists to prevent:

```{r, eval = has_both, error = TRUE}
compiled <- ffm_batch(jobs, run = FALSE, manifest = TRUE, .f = function(input, output, ...) {
  ffm_files(input, output) |> ffm_drop("video")
})

ffm_manifest(compiled)
```

## Bound a run that hangs

Every tidymedia call waits for the program it started to finish. A program that
hangs — a network path that stalls, a malformed file FFmpeg will not give up on
— takes the R session down with it, and in a batch of a thousand files that is
the whole run. Set a wall-clock limit and the hang stops the call instead.

The limit is a session option, in whole seconds:

```{r, eval = FALSE}
options(tidymedia.timeout = 600)
```

Unset, it is `0`, which means no limit — so nothing you already run changes
behavior until you ask for a bound.

To bound one call rather than the session, wrap it in `with_timeout()`. The
session's setting, or the absence of one, is back when the call returns, by any
route:

```{r, eval = has_both}
with_timeout(
  extract_audio("session01.mp4", "bounded.m4a"),
  seconds = 60
)

getOption("tidymedia.timeout")
```

To bound the rest of a function rather than an expression you wrap, say it as a
statement with `local_timeout()`. Every program started after that line is
bounded, and the caller's setting comes back when the function returns:

```{r, eval = has_both}
convert_all <- function(files) {
  local_timeout(60)
  vapply(files, function(f) extract_audio(f, sub("[.]mp4$", ".m4a", f)),
         character(1))
}

convert_all("session01.mp4")

getOption("tidymedia.timeout")
```

### What the limit actually bounds

The limit bounds **how long R waits**, not how long the program runs, and the
two are not the same number. R escalates: it interrupts the program at the
limit, asks it to terminate 20 seconds later, and kills it 20 seconds after
that. A program that ignores the first two signals is therefore waited for up to
**40 seconds longer than the limit you set** — measured on 2026-08-28 at 42.0 s
under a 2 s limit, on a Linux runner, with the same reading on a macOS host. Set
a limit meaning "give up somewhere around here", not "return at exactly this
second", and leave that headroom in anything downstream that depends on the call
returning.

A fractional limit is refused rather than rounded, because R truncates it toward
zero and `0` is the no-limit sentinel, so a rounded `0.5` would leave the call
unbounded. Every seam refuses it: `with_timeout()`, as below, and equally
`local_timeout()` and `options(tidymedia.timeout = 0.5)` when a call reads the
option:

```{r, error = TRUE}
with_timeout(extract_audio("session01.mp4", "x.m4a"), seconds = 0.5)
```

### A reached limit is never silent

Every call that can start one of these programs either aborts or warns when the
limit is reached; none of them quietly returns as though nothing happened.

It **aborts**, naming the program and the limit, from the task verbs,
`ffm_run()`, the escape hatches `ffmpeg()`, `ffprobe()` and `mediainfo()`, and
from `verify_media()` — a probe that never answered is not an answer of "no",
and absorbing it there would blame a good output for the probe's failure.

It **warns** where one hung file must not discard the rest of the work.
`probe_all()` and the `probe_*()` accessors, the MediaInfo readers and the
`get_*()` helpers give an `NA` row and one warning at the end of the call.
`ffm_batch()` and the `*_batch()` verbs mark the row `success = FALSE`, as for
any failed job, and warn once at the end saying how many jobs timed out. The
provenance manifest above warns that it could not read a version, and records
`NA` for it.

To act on either outcome in code, the abort carries the condition class
`tidymedia_timeout`; the version-probe and dropped-track warnings carry
`tidymedia_probe_timeout`, and the batch warning `tidymedia_batch_timeout`.

## Where to next

- `vignette("workflow")` — an end-to-end preprocessing pipeline.
- `vignette("batch")` — the jobs-tibble batch runner in depth.
- `vignette("metadata")` — reading each file's metadata as a tibble.
- `?tidymedia` — the package's option seams in one place.
