---
title: "A research preprocessing workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A research preprocessing workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidymedia)
```

This vignette walks a realistic end-to-end pipeline: preparing a set of recorded
**dyadic-interaction sessions** — two people in conversation, one camera each —
for behavioral coding and acoustic analysis. It is the kind of job tidymedia is
built for: the same handful of steps applied identically to every file, with a
saved command for each so the whole pipeline is reproducible.

Every step below uses `run = FALSE`, which **compiles** the FFmpeg command
without executing it, so you can read exactly what each verb would do. Drop that
argument to actually process the files. Chunks that describe operating over a
whole folder are shown but not evaluated; where a step *is* evaluated we use the
tiny sample clip that ships with the package:

```{r}
session <- system.file("extdata", "sample.mp4", package = "tidymedia")
```

Picture a study folder with one file per camera per session:

```{r, eval = FALSE}
recordings <- list.files("study/raw", pattern = "\\.mp4$", full.names = TRUE)
recordings
#> [1] "study/raw/session01_camA.mp4" "study/raw/session01_camB.mp4"
#> [2] "study/raw/session02_camA.mp4" "study/raw/session02_camB.mp4"
#> ...
```

## 1. Standardize the recordings

Cameras rarely agree on resolution, frame rate, or codec, and inconsistent
inputs make everything downstream harder. `standardize_video()` re-encodes a
file to a common target:

```{r}
standardize_video(
  session, "session01_camA_std.mp4",
  width = 1280, height = 720, fps = 30,
  run = FALSE
)
```

Because this runs the same way for every file, reach for the batch sibling to do
the whole folder from a jobs tibble (see `vignette("batch")` for the batch
model). Every task verb has a `*_batch()` companion like this one:

```{r, eval = FALSE}
jobs <- tibble::tibble(
  input  = recordings,
  output = sub("/raw/", "/std/", recordings)
)
standardize_video_batch(jobs, width = 1280, height = 720, fps = 30)
```

## 2. Prepare the audio

Acoustic and transcription tools want clean, consistently loud audio.
`normalize_audio()` brings a file to a target integrated loudness (here −23 LUFS,
the EBU R128 broadcast reference):

```{r}
normalize_audio(session, "session01_camA_norm.mp4",
                target_loudness = -23, run = FALSE)
```

To hand the speech to a transcription or acoustic-analysis tool, pull the audio
into a standalone file. `convert_audio()` transcodes to whatever the output
extension implies — a `.wav` for an automatic-speech-recognition pipeline:

```{r}
convert_audio(session, "session01_camA.wav", run = FALSE)
```

(If you only need the audio *without* re-encoding, `extract_audio()` stream-copies
it instead — see `vignette("tidymedia")`.)

## 3. Frames for visual coding

Frame-by-frame coding of facial behavior or gesture needs stills. `extract_frame()`
grabs a single frame at a timestamp — handy for a coding-manual illustration:

```{r}
extract_frame(session, "session01_camA_t30.png", timestamp = 30, run = FALSE)
```

`sample_frames()` samples at a fixed rate into a numbered image sequence — one
frame per second here, ready for a manual or computer-vision coding pass:

```{r, eval = FALSE}
sample_frames("session01_camA_std.mp4", outdir = "frames/session01_camA",
              fps = 1)
```

## 4. De-identify before sharing

Sharing recordings with remote coders usually means removing identifying
information first. Two verbs cover the common cases.

`strip_metadata()` removes container-level metadata — camera model, GPS
coordinates, recording timestamps — without touching the audio or video streams:

```{r}
strip_metadata(session, "session01_camA_clean.mp4", run = FALSE)
```

`anonymize_video()` box-fills one or more **fixed rectangles** you specify. This
is not face tracking — you give it the coordinates of a region that stays put,
such as an on-screen name caption or a fixed nameplate, and it paints a solid box
over it. Regions are a data frame, one row per box:

```{r}
regions <- tibble::tibble(
  x = 16, y = 640, width = 360, height = 64  # lower-left name caption
)
anonymize_video(session, "session01_camA_deid.mp4", regions = regions,
                run = FALSE)
```

## 5. Assemble and share

If a session was recorded in parts, `concatenate_videos()` joins them back into
one file (the parts must share a codec and container — they do here, since step 1
standardized them):

```{r, eval = FALSE}
concatenate_videos(
  c("session01_camA_part1.mp4", "session01_camA_part2.mp4"),
  "session01_camA_full.mp4"
)
```

Finally, `format_for_web()` produces a lightweight, broadly compatible H.264 file
with fast-start enabled, so coders can stream it in a browser without a large
download:

```{r}
format_for_web(session, "session01_camA_share.mp4", run = FALSE)
```

## Reproducibility

Every verb returns the exact FFmpeg command it compiled. Capturing those strings
— or running the pipeline through the batch runner, which collects them in a
`command` column — gives you a complete, re-runnable record of how each processed
file was produced. That record is the point: the pipeline above is not a
one-off, but a specification you can re-apply to the next cohort of recordings.

## Where to next

- `vignette("batch")` — the jobs-tibble batch runner in depth.
- `vignette("metadata")` — reading each file's metadata as a tibble.
- `vignette("tidymedia")` — the task verbs and the builder beneath them.
