---
title: "A research preprocessing workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A research preprocessing workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# The capability chunk in section 1 starts FFmpeg, so it is evaluated only when
# FFmpeg is installed. Every other evaluated chunk uses `run = FALSE` and starts
# nothing.
has_ffmpeg <- nzchar(Sys.which("ffmpeg"))
```

```{r setup}
library(tidymedia)
```

This page follows one research project from raw recordings to shared files.
The study records conversations between two people, with one camera on each
person. The recordings need the same steps before coding and sound analysis.

tidymedia suits this job. Each step runs the same way on every file, and each
step gives you a command that you can save and run again.

Most examples below use `run = FALSE`. The function then returns the FFmpeg
command without running it, so you can read what it would do. Leave out
`run = FALSE` to process the files.

Some examples work on a whole folder, so they are shown but not run. The others
use the short sample clip that comes with the package:

```{r}
session <- system.file("extdata", "sample.mp4", package = "tidymedia")
```

Say the study folder has one file for each camera in each session:

```{r, eval = FALSE}
jobs <- ffm_jobs("study/raw", type = "video")
jobs
#> # A tibble: 4 × 1
#>   input
#>   <chr>
#> 1 /data/study/raw/session01_camA.mp4
#> 2 /data/study/raw/session01_camB.mp4
#> 3 /data/study/raw/session02_camA.mp4
#> 4 /data/study/raw/session02_camB.mp4
```

`ffm_jobs()` returns a jobs table for the batch functions. It has one row for
each file, with the file's full path in an `input` column. The full path is why
the output above shows `/data/study/raw` and not `study/raw`.

If the folder has no video files, `ffm_jobs()` stops with an error. The next
section adds an `output` column to the table and passes it to a `*_batch()`
function.

## 1. Standardize the recordings

Cameras often differ in picture size, [frame rate](tidymedia.html#glossary) and
[codec](tidymedia.html#glossary). Those differences make later steps harder.
`standardize_video()` [re-encodes](tidymedia.html#glossary) a file to one
common format:

```{r}
standardize_video(
  session, "session01_camA_std.mp4",
  width = 1280, height = 720, fps = 30,
  run = FALSE
)
```

To do the same for the whole folder, use the batch version with the jobs table.
Each task function has a `*_batch()` version like this one. See
`vignette("batch")` for how batches work.

```{r, eval = FALSE}
jobs$output <- sub("/raw/", "/std/", jobs$input)
standardize_video_batch(jobs, width = 1280, height = 720, fps = 30)
```

### Using video hardware

Re-encoding a large study on the main processor can be slow. Functions that
re-encode video, such as `standardize_video()` and `format_for_web()`, take a
`hardware` argument. It moves the work to a
[hardware encoder](tidymedia.html#glossary).

There are two choices. Use `"nvenc"` for an NVIDIA graphics card, with an
FFmpeg build that supports nvenc. Use `"videotoolbox"` for Apple hardware on
macOS.

These functions show what your FFmpeg can do. `ffmpeg_codecs()` and
`ffmpeg_encoders()` list its codecs and encoders. `hardware_encoder()` gives
the encoder name for a codec and a hardware choice. `has_hardware_encoder()`
says whether your FFmpeg has that encoder:

```{r, eval = has_ffmpeg}
head(ffmpeg_codecs())

encoders <- ffmpeg_encoders()
head(encoders[encoders$type == "Video", c("name", "details")])

hardware_encoder("h264", "nvenc")
hardware_encoder("h264", "videotoolbox")
has_hardware_encoder("h264", "videotoolbox")
```

`has_hardware_encoder()` checks how FFmpeg was built. A `TRUE` result does not
prove that the hardware works. When it is `TRUE`, name that hardware in the
batch:

```{r, eval = FALSE}
standardize_video_batch(
  jobs, width = 1280, height = 720, fps = 30,
  hardware = "videotoolbox"
)
```

If the hardware you name is not available, the call stops with an error. So a
shared script never changes the codec without telling you. Add
`fallback = TRUE` to use the main processor instead.

The two choices support different codecs. NVIDIA nvenc supports H.264, HEVC
and AV1. Apple videotoolbox supports H.264 and HEVC. A codec that the hardware
does not support is an error that names both.

A hardware encoder trades some quality for speed. At the same bit rate, a
hardware encoder gives a picture that is a little worse than the software
encoder `libx264` or `libx265`. So it suits previews and bulk conversion, and
the software encoder suits a copy that you will keep or analyze frame by frame.

tidymedia does not do hardware decoding or run filters on a graphics card. For
those, use the direct command `ffmpeg()`.

## 2. Prepare the audio

Sound analysis and transcription tools work best with clean audio at an even
loudness. `normalize_audio()` sets a file's loudness to a target. Here the
target is -23 [LUFS](tidymedia.html#glossary), the level that the EBU R 128
broadcast standard uses.

The output of `normalize_audio()` has one audio
[stream](tidymedia.html#glossary) and no video. So give it an audio file name.
An `.mp4` name would give you a video file with sound and no picture:

```{r}
normalize_audio(session, "session01_camA_norm.wav",
                target_loudness = -23, run = FALSE)
```

To set the loudness and keep the picture, first write the audio file as above.
Then use the direct command `ffmpeg()` to put it back with the video.

A transcription tool needs the audio in its own file. `convert_audio()`
converts the audio to the format that the output extension names. Here that is
a `.wav` file for a speech recognition tool:

```{r}
convert_audio(session, "session01_camA.wav", run = FALSE)
```

If you need the audio as it is, `extract_audio()` copies it without converting
it. See `vignette("tidymedia")`.

## 3. Frames for visual coding

Coding facial expressions or gestures frame by frame needs still images.
`extract_frame()` saves one frame at a given time. This is useful for a picture
in a coding manual:

```{r}
extract_frame(session, "session01_camA_t30.png", timestamp = 30, run = FALSE)
```

`sample_frames()` saves frames at a fixed rate, as numbered image files. Here it
saves one frame per second, for coding by hand or by a computer vision tool:

```{r, eval = FALSE}
sample_frames("session01_camA_std.mp4", outdir = "frames/session01_camA",
              fps = 1)
```

## 4. De-identify before sharing

Before you share recordings with coders, you often must remove information
that identifies people. Two task functions cover the common cases.

`strip_metadata()` removes the file's metadata, such as the camera model, GPS
location and recording time. It does not change the audio or the video:

```{r}
strip_metadata(session, "session01_camA_clean.mp4", run = FALSE)
```

`anonymize_video()` fills one or more rectangles with a solid color. It does
not find or follow faces. You give the position of an area that does not move,
such as a name caption on screen. You give the areas as a data frame, with one
row for each box:

```{r}
regions <- tibble::tibble(
  x = 16, y = 640, width = 360, height = 64  # lower-left name caption
)
anonymize_video(session, "session01_camA_deid.mp4", regions = regions,
                run = FALSE)
```

## 5. Assemble and share

If a session was recorded in parts, `concatenate_videos()` joins the parts into
one file. The parts must have the same codec,
[container](tidymedia.html#glossary), picture size and frame rate. Here they
do, because step 1 made them the same:

```{r, eval = FALSE}
concatenate_videos(
  c("session01_camA_part1.mp4", "session01_camA_part2.mp4"),
  "session01_camA_full.mp4"
)
```

Last, `format_for_web()` makes an H.264 file that most browsers can play.
It moves the file's index to the start, so coders can watch it in a browser
before it has fully downloaded:

```{r}
format_for_web(session, "session01_camA_share.mp4", run = FALSE)
```

## Reproducibility

Each task function returns the FFmpeg command it used. A batch run also keeps
these commands, in a `command` column. Save the commands, and you have a full
record of how each file was made. You can then run the same steps on the next
group of recordings.

The command records what you asked for. A manifest records what happened: the
FFmpeg version, the time, and checksums of the files. Run the batch with
`manifest = TRUE` and read the manifest with `ffm_manifest()`.

To check that each output has the duration, size and codecs you asked for, use
`verify_media()`. `vignette("verification")` covers both, and also shows how to
stop a file that hangs.

## Where to next

- `vignette("batch")` explains the batch functions in more detail.
- `vignette("metadata")` shows how to read each file's metadata into a tibble.
- `vignette("verification")` shows how to check outputs, record how files were
  made and limit run time.
- `vignette("tidymedia")` explains the task functions and the pipeline
  functions.
